---
title: "Dplyr"
author: "Bharat Reddy"
date: "January 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Subset rows and filter columns

What should we do first when we open a new markdown

Load Libraries
```{r}
library(dplyr)
library(tidyverse)
```
Load Data
```{r}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim="\t") 
```

Goals:
subset rows of a dataframe and select columns of a data frame

```{r}
research <- filter(counts_raw, articleType=="Research Article")


```

Filter to research articles to 2006
```{r}
research_2006 <-  filter(research, year=="2006")

```

Filter to articles with a tweet
```{r}
research_2006_tweet <- filter(research_2006, backtweetsCount>0)
nrow(research_2006_tweet)


```
Research articles in 2006 and facebook

```{r}
research_2006_fb <- filter(research, year==2006, facebookCommentCount>0)

nrow(research_2006_fb)
```

Use logical operators
and &
or |

Look for research in 2006 with a tweet or facebook comment

```{r}
research_2006_fb_tweet <- filter(research, year==2006, facebookCommentCount>0 | backtweetsCount>0)
nrow(research_2006_fb_tweet)
```
Article in 2006 with social media coverage and is about infectious diseases
```{r}
research_2006_fb_tweet_disease <- filter(research, year==2006, facebookCommentCount>0 | backtweetsCount>0, grepl("Infectious Diseases", plosSubjectTags))
nrow(research_2006_fb_tweet_disease)
```
Subset columns of a dataframe with function subset
```{r}
colnames(research)
```
Subset the columns with the article information using the dplyr function select()

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
colnames(article_info)
```

New dataframe with columns doi:authorsCount
```{r}
article_info <- select(research, doi:authorsCount)
colnames(article_info)
```

create a metrics dataframe
```{r}
metrics <- select(research, contains("Count"))
colnames(metrics)
```
Exclude a column after contains
```{r}
metrics <- select(research, contains("Count"), -authorsCount)
colnames(metrics)
```
```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

Journal Column

```{r}
head(select(research, journal))

```

```{r}
head(select(research,3))
```
Slice: Slicing only the rows we want by numbers
Rows 1:3 article info dataframe
```{r}
slice(article_info, 1:3)
```

Reasearch before 2009
```{r}
research_before_2009 <- filter(research, year<2009)
research_before_2009_pdf <- filter(research, year<2009, pdfDownloadsCount>1000)
research_before_2009_pdf_men <- filter(research, year<2009, pdfDownloadsCount>1000, mendeleyReadersCount>15)
low_cite <- filter(research, year<2009, pdfDownloadsCount>1000, mendeleyReadersCount>15, wosCountThru2011<10)
nrow(low_cite)

```

#Channing commands with dplyr
In dplyer we use 

```{r}
facebook_2006<-research %>% filter(year==2006) %>% select(contains("facebook"))
head(facebook_2006)
```

```{r}
research %>% filter(year==2006) %>% select(contains("facebook")) %>% head
```

Arrange function: Use to sort rows in a dataframe

```{r}
research %>% arrange(authorsCount, wosCountThru2011) %>% select(authorsCount, wosCountThru2011) %>% slice(1:10)
```

flip the sort
```{r}
research %>% arrange(authorsCount, desc(wosCountThru2011)) %>% select(authorsCount, wosCountThru2011) %>% slice(1:10)
```

most cited 
```{r}
research %>% arrange(desc(wosCountThru2011)) %>% select(title) %>% slice(1:3)
```

lots of authors 
```{r}
research %>% arrange(desc(authorsCount)) %>% select(authorsCount, title, journal, plosSubjectTags) %>% slice(1:3)
```

#Summarize data with dplyr
Goals:
*new column size
*summarize with summarize
*count number of observations n()
*group data by variable (group_by)

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished/7)
research %>% select(weeksSincePublished) %>%  head()
```


summarize the data
```{r}
summarize(research, plos_mean=mean(plosCommentCount))
```

```{r}
plot_stats <- research %>% summarise(plos_mean=mean(plosCommentCount), plos_sd=sd(plosCommentCount))
plot_stats
```

```{r}
research %>%  filter(journal=="pone", year==2007) %>% 
  summarise(plos_mean=mean(plosCommentCount), plos_sd=sd(plosCommentCount))
```
add number of observations
```{r}
research %>%  filter(journal=="pone", year==2007) %>% 
  summarise(plos_mean=mean(plosCommentCount), plos_sd=sd(plosCommentCount), num=n())
```
Use group by to get stats by a grouping of the data
using BaseR
```{r}
mean_metric_per_var <- function(metric,variable){
  result=integer(length(variable)
  for (v in varaible){
    result[v] <- means(metric[variable==v])
  }
return(result)  
}
```

Using dplyr
```{r}
research %>%  group_by(journal) %>% summarise(tweets_mean=mean(backtweetsCount))
```

Group by 2 variables

```{r}
research %>%  group_by(journal, year) %>% summarise(tweets_mean=mean(backtweetsCount))

```
```{r}

tweets_per_journal <- research %>%  group_by(journal) %>% summarise(tweets_mean=mean(backtweetsCount), tweets_sem=sd(backtweetsCount)/sqrt(n()) )
tweets_per_journal
```
