---
title: "Trial Report"
author: "Arjun Biddanda"
date: "1/12/2019"
output: 
  html_document:
    toc : true
    number_sections : true
    highlight : espresso
    fig_caption : yes
    self_contained : false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial Analysis

## Reading in data on PLoS Papers

```{r read_data, results='hide', message=FALSE}
library(tidyverse)
counts_raw <- read_delim("data/counts-raw.txt.gz", delim = "\t")
n_pubs <- dim(counts_raw)[1]
```

## Diagnostic Plots

```{r histogram_weeks_since_pubs, fig.cap="Figure 1 : Weeks since publication per paper", fig.height=3, fig.width=5}
  weeks_since_pub <- counts_raw$daysSincePublished / 7
  hist(weeks_since_pub, xlab = "# of Weeks", main="Weeks since Publication")
```

```{r scatterplot_pubs_vs_citations}
plot(counts_raw$daysSincePublished, counts_raw$wosCountThru2011)
```

## Some vague statements about our data

Here is an example of a mathematical function : $y = x^2$

This dataset consists of `r n_pubs` publications.

## Exercise : 

  1. Plot a histogram of your favorite variable from the dataset
  2. make a scatter plot of your favorite variable against citation counts
    a. Comment on whether it has a marked effect on citation counts
    b. If appropriate create a scatter plot of log-transformed values
  3. Verify that the plots would be well-received by a collaborator!!

## Result

```{r}
hist(counts_raw$year, xlab="Year", main="Number of Publications per Year")
```


```{r}
plot(counts_raw$year,counts_raw$wosCountThru2011)
```


### A quick refresher on functions in `R`


```{r}


counts_raw[,"authorsCount"]
counts_raw[["authorsCount"]]


data <- counts_raw$authorsCount
plot(x = counts_raw$wosCountThru2011,y=data)

data_column <- "deliciousCount"
data <- counts_raw[[data_column]]
log_data <-log(data)
plot(x = counts_raw$wosCountThru2011,y=log_data)



```


```{r}
log_data <- function(input_dataframe,column){
  data_column <- input_dataframe[[column]]
  log_data_column <- log(data_column)
  return(log_data_column)
}

output <- log_data(input_dataframe = counts_raw,
         column = "deliciousCount")
plot(x = counts_raw$authorsCount,y=output)
```

```{r}
publication_date <-counts_raw$pubDate
head(publication_date)

num_unique <- function(x){
  unique_data <- unique(x)
  length_unique <- length(unique_data)
  return(length_unique)
}
debug(num_unique)
num_unique(x = counts_raw$pubDate)
```

```{r}
column_num_unique <- function(data,column_name){
  data_column <- data[[column_name]]
  if(is.null(data_column)){
    stop("No such column in data!")
  }
  unique_entries <- num_unique(data_column)
  return(unique_entries)
}
debug(column_num_unique)
column_num_unique(data = counts_raw,column_name = "doi")

```

### For loops and for fun


```{r}
fake_fun <- function(){
 x <- 4 
  
  
}
```


```{r}
names_of_columns <- colnames(counts_raw)
my_column <- "doi"
column_num_unique(data = counts_raw,
                  column_name = my_column)

unique_col_ct <- function(data){
  data_column_names <- colnames(data)
  output_length <- length(data_column_names)
  result <- integer(output_length)
  for(i in seq_along(data_column_names)){
    my_column <- data_column_names[i]
    result[i] <- column_num_unique(data,my_column)
  }
  print("we finished!")
  return(result)
}

debug(unique_col_ct)
doy_cols <- which(colnames(counts_raw)=="doy")
unique_col_ct(counts_raw[,doy_cols])
```


```{r}
counts_columns <- c("facebookShareCount",
                    "facebookLikeCount")

max_of_each_col <- function(data,count_colnames){
  number_of_columns <- length(count_colnames)
  result <- integer(number_of_columns)
  for(i in seq_along(count_colnames)){
    my_column <- count_colnames[i]
    my_vector <- data[[my_column]]
    result[i] <- max(my_vector)
  }
  return(result)
}

debug(max_of_each_col)
max_of_each_col(counts_raw,counts_columns)

```


## Quick refresher on subsetting with R

There are many ways to subset data in R
### We can subset a vecter by position with `[`
```{r}
x <- c(8,9,3,5,7)
x[4]
y <- c(1,2)
x[y]
z <- length(x)
x[z]
```
We can also subset using double bracket syntax
to get *one* element back
```{r}
x[[1]]
x[[y]]
```
## Indexing dataframes
Dataframes can also be indexed with single bracket `[` and double bracket `[[` syntax depending on what we want

```{r}
counts_raw[,1]
y <- c(1,2)
counts_raw[,y]
counts_raw[,c("doi","daysSincePublished")]
(counts_vector <- counts_raw[["doi"]])
(doi_dataframe <- counts_raw[,"doi"])
```







